Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A guide for Vietnamese and International Students

### 2023-06-14 phd-cs-us.tex
updates

\begin{commentbox}[Thanh]
Vietnamese universities typically offer several specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year). However, these programs may present several challenges. Professors within these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings. For example, an 3.5 GPA students from such talented programs are easily much better than a 4.0 GPA students not in those programs.  Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities. All of these make gaining admission in the US difficult as US faculty are not familiar with these issues. I wonder if you have and advice and tips for the situations? 
\tcblower
\textbf{Vu} Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad. One way to mitigate this is making these issues known to admission committee.  Schools with Vietnamese profs are probably aware of them, but in general your letter writers/you can explicitly mention these in their letters/ your statement.
\end{commentbox}

@@

As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
RA is the type of funding paid by a prof. for you to do their research. TA is paid by the department for you to help with teaching. Finally, fellowship is an independent funding that can come from a school, a company, or an organization. Note that funding is typically more available for PhD students than 
Masters and especially undergraduate studies, which typically have no funding.  Tab.~\ref{tab:funding} summarizes the differences.

>>>>

\begin{commentbox}[Thanh]
Vietnamese universities typically offer several specialized programs, such as the talented engineer program at HUST, that have highly competitive entrance exams and a limited number of available slots (e.g., 30 per year). However, these programs may present several challenges. Professors within these programs often set higher requirements for students, including more demanding tests and assignments, resulting in lower GPAs and overall rankings. For example, an 3.5 GPA students from such talented programs are easily much better than a 4.0 GPA students not in those programs.  Similarly, variations in GPA standards exist among different universities, with technical universities generally having lower GPAs than economical universities. All of these make gaining admission in the US difficult as US faculty are not familiar with these issues.
\tcblower
\textbf{Vu} Many Vietnamese students and faculty lament how this competitive/low-grading from top Vietnamese universities hurts its own students in applying abroad. One way to mitigate this is making these issues known to admission committee.  Schools with Vietnamese profs are probably aware of them, but in general your letter writers/you can explicitly mention these in their letters/ your statement.
\end{commentbox}

@@

As mentioned, if you're admitted to a \emph{good} CS PhD program, you should not have to worry about funding!  
In the US, the common types of funding for PhD are \emph{graduate teaching assistant} (GTA or TA), \emph{graduate research assistant} (GRA or RA), and \emph{Fellowship}.
RA is paid by a prof. for you to do their research. TA is paid by the department for you to help with teaching. Finally, fellowship is an independent funding that come from a school, a company, or an organization. Tab.~\ref{tab:funding} summarizes the differences.

Note that funding is typically more available for PhD students than 
Masters and especially undergraduate studies, which typically have no funding (and they also have to pay international tuition, which is very expensive!).
